### 2004-07-19 Documentation/part1.tex
Editing done ;-)

\chapter{A \GLOBES\ tour}
\label{chapter:tour}
\index{norm}{GLoBES tour}

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

@@

\index{norm}{Clear experiment list}
\index{norm}{Delete experiments}
One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{ClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{\_params} and \GLB{\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

@@

Similarly, the density parameters or number of iterations 
(returned by the minimizers) can be accessed:
\begin{function}
\GLBNS{SetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set accordingly.
\end{function}
\begin{function}
\GLBNS{GetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function} 
\begin{function}
\GLBNS{SetIteration} \index{norm}{Iterations}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:\index{norm}{Parameter vector handling}
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors freed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
\index{norm}{Systematics $\chi^2$}

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}
\index{norm}{Two-parameter correlation}
A typical and fast application for \GLB{ChiSys} is the visualization of
 two-parameter correlations using systematics only. For example, to
  calculate the two-parameter 
 correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
 can use the following code excerpt from {\tt example1.c}:
\begin{quote}
{\tt {\footnotesize
  /* Initialize parameter vector(s) and compute simulated data */ \\
  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
  glbSetOscillationParameters(true\_values); glbSetRates(); \\
\\
  /* Iteration over all values to be computed */ \\
  double x,y,res; \\   
  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
  \{ \\
\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
  \\
\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
\hspace*{0.5cm} AddToOutput(x,y,res); \\
  \} \\
} }
\end{quote}

@@

\index{norm}{Systematics} \index{norm}{Pull method}
The treatment of systematics in \GLOBES\ is performed by the so-call
{\em pull method} with the help of auxiliary systematics parameters.
They are taken completely uncorrelated among different rules, and treated with simple Gaussian statistics. In general, a rule is a set 
of signal and background event rates coming from different oscillation
channels, where the event rates of all rule contributions are added.
For more details of the rule concept, see \Part~\ref{part:2} of this manual,
and for the treatment of systematics, see \Sec~\ref{sec:rules}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Calculating $\chi^2$-projections: how one can include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how one can include correlations}
\label{chapt:correlations}
\index{norm}{Correlations $\chi^2$}

\index{norm}{Correlations!multi-parameter}
This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
to include the full $n$-parameter correlation in the high-dimensional parameter space with reasonable effort. Of course, calculating $\chi^2$-projections is somewhat more complicated than using systematics only. Therefore, we use a simple step by step introduction to the problem.

@@

\index{norm}{Priors} 
\index{norm}{Input errors} 
\index{norm}{Starting values}
In many cases, the fit manifold is restricted by the knowledge 
from earlier experiments. For example, the knowledge on the solar 
parameters will in most cases be supplied by the solar neutrino experiments. 
If the external precision of a parameter is at the time of the 
measurement better than the one of the experiment itself, one usually will use
this external, better knowledge and impose a corresponding constraint
 on this parameter. This external knowledge may reduce the extension of the 
$n$-dimensional fit manifold in the respective direction. In the most 
extreme case, keeping all parameters but the measured one fixed 
in the analysis is equivalent to the assumption that all parameters 
are determined externally with infinitively high precisions. As this 
is a quite strong assumption, one should \emph{always} check the 
consequences of relaxing it and using realistic errors. Only if such
a test has demonstrated that the impact of the uncertainty on a 
given fit parameter is negligible it can be assumed as fixed safely. 
 The inclusion of external input in \GLOBES\ is done by the use of 
Gaussian {\em priors}: We assume that an external measurement has 
determined the measured parameter to be at the central value 
(which we call {\em starting value}) with a $1 \sigma$ Gaussian error 
(which we call {\em input error}). The explicit definition of these 
priors will be shown in the next section.

@@

\index{norm}{External input}
It is one of the strengths of the \GLOBES\ software to use external input 
in order to reduce the extension of the fit manifold with the knowledge 
from external
(earlier) measurements. The treatment of external input is done by the 
addition of Gaussian so-called {\em priors} to the systematics-minimized 
$\chi^2$-function. For example, for the matter density, one obtains as the 
final projected $\chi^2_F$ after minimization over the matter density
scaling factor \index{norm}{Matter density!scaling factor} $\hat{\rho}$
\be
 \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) +
 \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
\label{equ:priors}
\ee
This example is a very simple one, since in fact the
minimization is simultaneously performed over all priors and free oscillation 
parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the 
prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) 
{\em input error}. Thus, it is assumed that an external measurement has 
determined the matter density with a precision (input error) 
$\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$. Usually, 
the starting value is fixed at the best-fit value, and the input error 
to the $1 \sigma$ half width of the external measurement. For the matter 
density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
the actual matter density profile such as given by the experiment
 definition file, and $\sigma_{\hat{\rho}}$ to the 
relative matter density uncertainty (\eg, $0.05$ for 5\% uncertainty).

@@

\index{norm}{Input errors} 
\index{norm}{Starting values}
In order to set the starting values and input errors, two function have to
 be called {\em before the usage of any minimizer}:
\begin{function}
\GLBNS{SetStartingValues}
{\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
\end{function}
\begin{function}
\GLBNS{SetInputErrors}
{\tt int glbSetInputErrors(const glb\_params in)}
 sets the input errors for all of the following minimizer calls to {\tt in}.
 An input error of $0$ corresponds to not taking into account the
 respective prior.
\end{function}
Accordingly, there are functions to return the actually set starting values
and input errors:
\begin{function}
\GLBNS{GetStartingValues}
{\tt int glbGetStartingValues(glb\_params out)} writes the currently
set starting values to {\tt out}.
\end{function}
\begin{function}
\GLBNS{GetInputErrors}
{\tt int glbGetInputErrors(glb\_params out)}
 writes the currently set input errors to {\tt out}.
\end{function}
All functions take or return as many matter density parameters as there are initialized experiments. In addition, they return $-1$ if the operation
was not successful.

\index{norm}{External precision}
Eventually, a typical initialization of the external input with
$10\%$ external precisions for the solar 
parameters\footnote{In fact, accelerator-based long-baseline experiments 
are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, 
which means that these errors effectively add up to an error of this 
product of about $15\%$.}, 
and $5\%$ matter density uncertainties for all experiments looks like this:
\begin{quote}
{\tt
 glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
 glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
 glbSetStartingValues(true\_values);\\
 glbSetInputErrors(input\_errors);\\
}
\end{quote}
In this example, the starting values are set to 
the true (simulated) values. Remember that initially the 
matter density scaling factors are 
all $1.0$, which means that they do not need to be adjusted for the
starting values.

\index{norm}{Priors}
\index{norm}{Advanced tricks}
Though the priors are an elegant way to treat external input, 
there are also some complications with priors. The following hints are for the
more advanced \GLOBES\ user:
\begin{enumerate}
\item
 The priors are only added once to the final $\chi^2$, no matter how
 many experiments there are simulated. This is already one reason
 (besides the minimization) why the sum of all projected $\chi^2$'s of 
 the individual experiments 
 cannot correspond to the $\chi^2$ of the combination of all experiments.
\item
 Priors are not used for parameters which are not minimized over, \ie, 
 kept fixed. This will be important together with arbitrary projections
 using \GLB{ChiNP}. A more subtile consequence is the comparison
 of fit manifold sections and projections for the solutions where
 the absolute minimum $\chi^2$ is larger than zero, \ie, degeneracies
 other than the true solution. In this case, the sections and
 projections are not  comparable if not corrected by the prior contributions, where the
 correction can be obtained as the $\chi^2$-difference at the minimum.
 For example, projecting the $\mathrm{sgn}(\ldm)$-degeneracy onto
 the $\theta_{13}$-$\deltacp$-plane and comparing it with the section
 (all other parameters fixed), the section region would in many cases be 
 larger than the projection region if the priors are not added to the
 section. At the true solution, this problem usually does not occur
 because the prior contributions are close to zero.    
\item
Currently, \GLOBES\ only supports Gaussian priors for the individual
oscillation parameters. Especially for the solar parameters, this is only
 an approximation, since they are imposed on $\theta_{12}$ and not
 on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
 \theta_{12}$. Later versions of  \GLOBES\ may include more alternatives.
 \end{enumerate}

@@

\begin{table}[t!]
\begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
\hline
Function & Purpose & Parameters $\rightarrow$ Result\\
\hline
\GLB{AllocProjection} & Allocate projection vector & {\tt ()} \\
\GLB{FreeProjection} & Free projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
\GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
\GLB{CopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
\GLB{PrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
\GLB{SetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
\GLB{GetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
 {\tt glbSetDensity\-ProjectionFlag} \GLBNS{SetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
{\tt glbGetDensity\-ProjectionFlag} \GLBNS{GetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.  & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
\hline
\end{tabular}
\caption{\label{tab:defprojection}\index{norm}{Projection type}
Different functions handling the
\GLB{\_projection} type. Flags are either \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}. The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
\end{table}

@@

\index{norm}{Projection: Define}
As soon as we have defined a projection, we can assign it:
\begin{function}
\GLBNS{SetProjection}
{\tt int glbSetProjection(const glb\_projection in)} sets the projection
to {\tt in}. The return value is $0$ if successful, and $-1$ if
unsuccessful.
\end{function}
Similarly, the currently assigned projection can be returned with:
\begin{function}
\GLBNS{GetProjection}
{\tt int glbGetProjection(glb\_projection out)} writes the currently
set projection to {\tt out}. The return value is $0$ if successful, and $-1$ if unsuccessful.
\end{function}
After setting the starting values, input errors, and the projection, 
we can run the minimizer:
\begin{function}
\GLBNS{ChiNP}
{\tt double glbChiNP(const glb\_params in, glb\_params out, int exp)} 
returns the projected $\chi^2$ onto the hyperplane specified by 
{\tt glbSetProjection} for the  experiment {\tt exp}. 
For the simulation of all initialized experiments,
use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed fit values for the minimizer (all free parameters) and the fit values
on the hyperplane (all fixed parameters). The actually determined parameters at the minimum are returned in {\tt out}, where the fixed parameters are still at their input values. If {\tt out} is set to {\tt NULL}, this information will not be returned.
\end{function}
As an example, the projection sequence for a minimization over
$\deltacp$ only looks like this:
\begin{quote}
{\tt
  glb\_projection th13\_projection = glbAllocProjection(); \\
  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,\\
  \hspace*{2cm} GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
  glbSetProjection(th13\_projection); \\ 
  res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
  glbFreeProjection(th13\_projection);
}
\end{quote}
In this case, only the correlation with $\deltacp$ is taken into account.
Note that in the  example on page~\pageref{ex:corrproj} this projection
is compared with the result including the full multi-parameter correlation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Locating degenerate solutions}
\index{norm}{Degenerate solutions}
\index{norm}{Degeneracies $\chi^2$}

@@

\example{Finding the $\mathrm{sgn}(\ldm)$-degeneracy}{
\label{ex:sgndeg}
\index{norm}{Degeneracies: sgn($\ldm$)-degeneracy}

@@

\section{Systematics}
\label{sec:systematics}
\index{norm}{Systematics}
\index{norm}{Systematics on/off}

@@

\section{Algorithm parameters: Filter functions}
\index{norm}{Filter functions}

>>>>

\chapter{A \GLOBES\ tour}
\label{chapter:tour}
\index{norm}{GLOBES tour@\GLOBES~tour}

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{glb files@{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors.}
\end{table}

@@

\index{norm}{Experiment!list!clear}
\index{norm}{Experiment!delete}
One can also remove all experiments from the evaluation list at running
 time:
\begin{function}
\GLBNS{ClearExperimentList}
{\tt void glbClearExperimentList()} removes all experiments from the internal list and resets all counters.   
\end{function}
Note that changing the number of experiments requires a new initialization
of all parameters of the types \GLB{\_params} and \GLB{\_projection}
if the number of experiments changes, since these parameter structures internally carry lists for the matter densities of all experiments. Similarly, once should never call {\tt glbAlloc...} before the
experiment initialization.

@@

Similarly, the density parameters or number of iterations 
(returned by the minimizers) can be accessed:
\begin{function}
\GLBNS{SetDensityParams}
{\tt glb\_params glbSetDensityParams(glb\_params in, double dens, int which)} sets the density parameter {\tt which} in the structure {\tt in}
to the value {\tt dens}. If the assignment was unsuccessful, the 
function returns {\tt NULL}. If {\tt GLB\_ALL} is used for {\tt which},
the density parameters of all experiments will be set accordingly.
\end{function}
\begin{function}
\GLBNS{GetDensityParams}
{\tt double glbGetDensityParams(glb\_params in, int which)}
returns the value of the density parameter {\tt which} in the 
structure {\tt in}.
\end{function} 
\begin{function}
\GLBNS{SetIteration} \index{norm}{Minimizer!iterations}
{\tt glb\_params glbSetIteration(glb\_params in, int iter)} sets the number of iterations in the structure {\tt in}
to the value {\tt iter}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetIteration}
{\tt int glbGetIteration(glb\_params in)}
returns the value of the number of iterations in the 
structure {\tt in}.
\end{function}
In total, the parameter vector handling in a program normally has the
following order:\index{norm}{Parameter vector handling}
\begin{quote}
{\tt
glbInitExperiment(...); \\
/* ... more initializations ... */ \\
\\
glb\_params vector1 = glbAllocParams(); \\
/* ... more vectors allocated ... */ \\
\\
/* Program code: assign and use vectors */ \\
\\
glbFreeParams(vector1); \\
/* ... more vectors freed ... */ \\
\\
/* ... end of program or glbClearExperimentList ... */ 
}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Calculating $\chi^2$ with systematics only]{Calculating $\boldsymbol{\chi^2}$ with systematics only}
\index{norm}{Systematics!$\chi^2$}

@@

\example{Correlation between $\stheta$ and $\deltacp$}{
\label{ex:corrth13dcp}
\index{norm}{Correlation!two-parameter}
A typical and fast application for \GLB{ChiSys} is the visualization of
 two-parameter correlations using systematics only. For example, to
  calculate the two-parameter 
 correlation between $\stheta$ and $\deltacp$ at a neutrino factory, one
 can use the following code excerpt from {\tt example1.c}:
\begin{quote}
{\tt {\footnotesize
  /* Initialize parameter vector(s) and compute simulated data */ \\
  glbDefineParams(true\_values,theta12,theta13,theta23,deltacp,sdm,ldm); \\
  glbDefineParams(test\_values,theta12,theta13,theta23,deltacp,sdm,ldm);  \\
  glbSetOscillationParameters(true\_values); glbSetRates(); \\
\\
  /* Iteration over all values to be computed */ \\
  double x,y,res; \\   
  for(x=-4.0;x<-2.0+0.01;x=x+2.0/50) \\
  for(y=0.0;y<200.0+0.01;y=y+200.0/50) \\
  \{ \\
\hspace*{0.5cm} /* Set parameters in vector of test values */ \\
\hspace*{0.5cm} glbSetOscParams(test\_values,asin(sqrt(pow(10,x)))/2,GLB\_THETA\_13); \\
\hspace*{0.5cm} glbSetOscParams(test\_values,y*M\_PI/180.0,GLB\_DELTA\_CP); \\
  \\
\hspace*{0.5cm} /* Compute Chi2 for all loaded experiments and all rules */ \\
\hspace*{0.5cm} res=glbChiSys(test\_values,GLB\_ALL,GLB\_ALL); \\
\hspace*{0.5cm} AddToOutput(x,y,res); \\
  \} \\
} }
\end{quote}

@@

\index{norm}{Systematics} \index{norm}{Pull method}
The treatment of systematics in \GLOBES\ is performed by the so-call
{\em pull method} with the help of auxiliary systematics parameters.
They are taken completely uncorrelated among different rules, and treated with simple Gau\ss ian statistics. In general, a rule is a set 
of signal and background event rates coming from different oscillation
channels, where the event rates of all rule contributions are added.
For more details of the rule concept, see \Part~\ref{part:2} of this manual,
and for the treatment of systematics, see \Sec~\ref{sec:rules}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Calculating $\chi^2$-projections: how one can include correlations]{Calculating $\boldsymbol{\chi^2}$-projections: how one can include correlations}
\label{chapt:correlations}
\index{norm}{Correlation!and $\Delta\chi^2$}

\index{norm}{Correlation!multi-parameter}
This chapter deals with the rather complicated issue of $n$-parameter correlations. It is one of the greatest strengths of this software 
to include the full $n$-parameter correlation in the high-dimensional parameter space with reasonable effort. Of course, calculating $\chi^2$-projections is somewhat more complicated than using systematics only. Therefore, we use a simple step by step introduction to the problem.

@@

\index{norm}{External information!priors} 
\index{norm}{External information!input errors} 
\index{norm}{External information!starting values}
In many cases, the fit manifold is restricted by the knowledge 
from earlier experiments. For example, the knowledge on the solar 
parameters will in most cases be supplied by the solar neutrino experiments. 
If the external precision of a parameter is at the time of the 
measurement better than the one of the experiment itself, one usually will use
this external, better knowledge and impose a corresponding constraint
 on this parameter. This external knowledge may reduce the extension of the 
$n$-dimensional fit manifold in the respective direction. In the most 
extreme case, keeping all parameters but the measured one fixed 
in the analysis is equivalent to the assumption that all parameters 
are determined externally with infinitively high precisions. As this 
is a quite strong assumption, one should \emph{always} check the 
consequences of relaxing it and using realistic errors. Only if such
a test has demonstrated that the impact of the uncertainty on a 
given fit parameter is negligible it can be assumed as fixed safely. 
 The inclusion of external input in \GLOBES\ is done by the use of 
Gau\ss ian {\em priors}: We assume that an external measurement has 
determined the measured parameter to be at the central value 
(which we call {\em starting value}) with a $1 \sigma$ Gau\ss ian error 
(which we call {\em input error}). The explicit definition of these 
priors will be shown in the next section.

@@

\index{norm}{External information}
\index{norm}{External input| \see{External information}}
It is one of the strengths of the \GLOBES\ software to use external input 
in order to reduce the extension of the fit manifold with the knowledge 
from external
(earlier) measurements. The treatment of external input is done by the 
addition of Gau\ss ian so-called {\em priors} to the systematics-minimized 
$\chi^2$-function. For example, for the matter density, one obtains as the 
final projected $\chi^2_F$ after minimization over the matter density
scaling factor \index{norm}{Matter density!scaling factor} $\hat{\rho}$
\be
 \chi^2_F = \underset{\rho}{\mathrm{min}} \left( \chi^2(\hat{\rho}) +
 \frac{(\hat{\rho} - \hat{\rho}^0)^2}{\sigma_{\hat{\rho}}^2} \right).
\label{equ:priors}
\ee
This example is a very simple one, since in fact the
minimization is simultaneously performed over all priors and free oscillation 
parameters. In \equ{priors}, $\hat{\rho}^0$ is the {\em starting value} of the 
prior, and $\sigma_{\hat{\rho}}$ the $1 \sigma$ absolute (half width) 
{\em input error}. Thus, it is assumed that an external measurement has 
determined the matter density with a precision (input error) 
$\sigma_{\hat{\rho}}$ at the central value $\hat{\rho}^0$. Usually, 
the starting value is fixed at the best-fit value, and the input error 
to the $1 \sigma$ half width of the external measurement. For the matter 
density, $\hat{\rho}^0$  is usually set to $1.0$ corresponding to
the actual matter density profile such as given by the experiment
 definition file, and $\sigma_{\hat{\rho}}$ to the 
relative matter density uncertainty (\eg, $0.05$ for 5\% uncertainty).

@@

\index{norm}{External information!input errors} 
\index{norm}{External information!starting values}
In order to set the starting values and input errors, two function have to
 be called {\em before the usage of any minimizer}:
\begin{function}
\GLBNS{SetStartingValues}
{\tt int glbSetStartingValues(const glb\_params in)} sets the starting values for all of the following minimizer calls to {\tt in}.
\end{function}
\begin{function}
\GLBNS{SetInputErrors}
{\tt int glbSetInputErrors(const glb\_params in)}
 sets the input errors for all of the following minimizer calls to {\tt in}.
 An input error of $0$ corresponds to not taking into account the
 respective prior.
\end{function}
Accordingly, there are functions to return the actually set starting values
and input errors:
\begin{function}
\GLBNS{GetStartingValues}
{\tt int glbGetStartingValues(glb\_params out)} writes the currently
set starting values to {\tt out}.
\end{function}
\begin{function}
\GLBNS{GetInputErrors}
{\tt int glbGetInputErrors(glb\_params out)}
 writes the currently set input errors to {\tt out}.
\end{function}
All functions take or return as many matter density parameters as there are initialized experiments. In addition, they return $-1$ if the operation
was not successful.

\index{norm}{External information!precision}
Eventually, a typical initialization of the external input with
$10\%$ external precisions for the solar 
parameters\footnote{In fact, accelerator-based long-baseline experiments 
are primarily sensitive to the product $\sin 2 \theta_{12} \cdot \sdm$, 
which means that these errors effectively add up to an error of this 
product of about $15\%$.}, 
and $5\%$ matter density uncertainties for all experiments looks like this:
\begin{quote}
{\tt
 glbDefineParams(input\_errors,theta12*0.1,0,0,0,sdm*0.1,0);\\  
 glbSetDensityParams(input\_errors,0.05,GLB\_ALL);\\
 glbSetStartingValues(true\_values);\\
 glbSetInputErrors(input\_errors);\\
}
\end{quote}
In this example, the starting values are set to 
the true (simulated) values. Remember that initially the 
matter density scaling factors are 
all $1.0$, which means that they do not need to be adjusted for the
starting values.

\index{norm}{Minimizer!priors}
\index{norm}{External information!priors}
\index{norm}{Advanced tricks}
Though the priors are an elegant way to treat external input, 
there are also some complications with priors. The following hints are for the
more advanced \GLOBES\ user:
\begin{enumerate}
\item
 The priors are only added once to the final $\chi^2$, no matter how
 many experiments there are simulated. This is already one reason
 (besides the minimization) why the sum of all projected $\chi^2$'s of 
 the individual experiments 
 cannot correspond to the $\chi^2$ of the combination of all experiments.
\item
 Priors are not used for parameters which are not minimized over, \ie, 
 kept fixed. This will be important together with arbitrary projections
 using \GLB{ChiNP}. A more subtile consequence is the comparison
 of fit manifold sections and projections for the solutions where
 the absolute minimum $\chi^2$ is larger than zero, \ie, degeneracies
 other than the true solution. In this case, the sections and
 projections are not  comparable if not corrected by the prior contributions, where the
 correction can be obtained as the $\chi^2$-difference at the minimum.
 For example, projecting the $\mathrm{sgn}(\ldm)$-degeneracy onto
 the $\theta_{13}$-$\deltacp$-plane and comparing it with the section
 (all other parameters fixed), the section region would in many cases be 
 larger than the projection region if the priors are not added to the
 section. At the true solution, this problem usually does not occur
 because the prior contributions are close to zero.    
\item
Currently, \GLOBES\ only supports Gau\ss ian priors for the individual
oscillation parameters. Especially for the solar parameters, this is only
 an approximation, since they are imposed on $\theta_{12}$ and not
 on $\sin 2 \theta_{12}$, $\sin 2 \theta_{12} \cdot \sdm$, or $\sin  
 \theta_{12}$. Later versions of  \GLOBES\ may include more alternatives.
 \end{enumerate}

@@

\begin{table}[t!]
\begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
\hline
Function & Purpose & Parameters $\rightarrow$ Result\\
\hline
\GLB{AllocProjection} & Allocate projection vector & {\tt ()} \\
\GLB{FreeProjection} & Free projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
\GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
\GLB{CopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
\GLB{PrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
\GLB{SetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
\GLB{GetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
 {\tt glbSetDensity\-ProjectionFlag} \GLBNS{SetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
{\tt glbGetDensity\-ProjectionFlag} \GLBNS{GetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.  & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
\hline
\end{tabular}
\caption{\label{tab:defprojection}\index{norm}{Projection!type}
Different functions handling the
\GLB{\_projection} type. Flags are either \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}. The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
\end{table}

@@

\index{norm}{Projection! definition}
As soon as we have defined a projection, we can assign it:
\begin{function}
\GLBNS{SetProjection}
{\tt int glbSetProjection(const glb\_projection in)} sets the projection
to {\tt in}. The return value is $0$ if successful, and $-1$ if
unsuccessful.
\end{function}
Similarly, the currently assigned projection can be returned with:
\begin{function}
\GLBNS{GetProjection}
{\tt int glbGetProjection(glb\_projection out)} writes the currently
set projection to {\tt out}. The return value is $0$ if successful, and $-1$ if unsuccessful.
\end{function}
After setting the starting values, input errors, and the projection, 
we can run the minimizer:
\begin{function}
\GLBNS{ChiNP}
{\tt double glbChiNP(const glb\_params in, glb\_params out, int exp)} 
returns the projected $\chi^2$ onto the hyperplane specified by 
{\tt glbSetProjection} for the  experiment {\tt exp}. 
For the simulation of all initialized experiments,
use {\tt GLB\_ALL} for {\tt exp}. The values in {\tt in} are the guessed fit values for the minimizer (all free parameters) and the fit values
on the hyperplane (all fixed parameters). The actually determined parameters at the minimum are returned in {\tt out}, where the fixed parameters are still at their input values. If {\tt out} is set to {\tt NULL}, this information will not be returned.
\end{function}
As an example, the projection sequence for a minimization over
$\deltacp$ only looks like this:
\begin{quote}
{\tt
  glb\_projection th13\_projection = glbAllocProjection(); \\
  glbDefineProjection(th13\_projection,GLB\_FIXED,GLB\_FIXED,GLB\_FIXED,\\
  \hspace*{2cm} GLB\_FREE,GLB\_FIXED,GLB\_FIXED); \\
  glbSetProjection(th13\_projection); \\ 
  res1=glbChiNP(test\_values,NULL,GLB\_ALL); \\
  glbFreeProjection(th13\_projection);
}
\end{quote}
In this case, only the correlation with $\deltacp$ is taken into account.
Note that in the  example on page~\pageref{ex:corrproj} this projection
is compared with the result including the full multi-parameter correlation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Locating degenerate solutions}
\index{norm}{Degeneracies|(}
\index{norm}{Degeneracies!multiple solutions}
\index{norm}{Degeneracies!and $\Delta\chi^2$}

@@

\example{Finding the $\mathrm{sgn}(\ldm)$-degeneracy}{
\label{ex:sgndeg}
\index{norm}{Degeneracies!sgn($\ldm$)-degeneracy}

@@

\section{Systematics}
\label{sec:systematics}
\index{norm}{Systematics}
\index{norm}{Systematics!on/off}

@@

\section{Algorithm parameters: Filter functions}
\index{norm}{Filter!functions}
